Predicting Solid State Material Platforms for Quantum Technologies

### 2022-01-17 paper_v2.tex
Updates from Overleaf

\begin{abstract}

%Rewrite to avoid plagiarism complaints, original is commented out

%Semiconductor materials provide a compelling platform for quantum technology, and a vast amount of materials and their properties can be found in high-throughput databases. However, sifting through tens of thousands of these materials in order to find novel candidates for quantum technology is a challenge. Therefore, we provide a framework for the automatic discovery of promising solid-state material hosts using machine learning methods.
Semiconductor materials provide a compelling platform for quantum technologies (QT), and the properties of a vast amount of materials can be found in databases containing information from both experimental and theoretical explorations. However, scouring these databases to find promising candidate materials for quantum technology applications among tens of thousands of contenders is a major challenge. Therefore, we have developed a framework for the automated discovery of semiconductor host platforms for QT using material informatics and machine learning methods.

%We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a data set consisting of more than $25000$ materials. Furthermore, we have developed and implemented three data mining approaches, termed \textit{the Ferrenti approach}, \textit{the augmented Ferrenti approach} and \textit{the intuitive approach} for defining three distinct training sets for the supervised machine learning algorithms logistic regression, decision tree, random forest and gradient boost to be trained on.
Data extraction tools have been developed for a number of materials science databases resulting in a data set consisting of over $25.000$ materials. The information from the databases was converted into nearly $5000$ physics-informed features. Furthermore, we have developed and implemented three different approaches for the data mining process: \textit{The Ferrenti approach}, \textit{the augmented Ferrenti approach}, and \textit{the intuitive approach}. 
%Rename insightful approach! To what? Restricted? Experience-based?  
%Options include stringent, strict, cherry-picked, informed and intuitive. 
The three approaches define distinct training sets for the supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting.

%We find a lack of consistent results for the Ferrenti approach and the augmented Ferrenti approach due to an overly broad formulation of the training set, whereas the restrictions set in the intuitive approach proved suitable. All models for the intuitive approach agreed on $214$ predicted candidates, with examples such as ZnGeP$_2$, MgSe, CdS, BP, BC2N, BP, Ge, GeC, InP, and InAs. All three approaches and all ML models agreed on a subset of $47$ eligible candidates of $8$ elemental, $29$ binary, and $10$ tertiary compounds.
%We find a lack of consistent results for the Ferrenti approach and the augmented Ferrenti approach due to an overly broad formulation of the training set. The restrictions set in the intuitive approach, which relies exclusively on findings from the literature, proved on the other hand more suitable.
We find that the Ferrenti approach and the augmented Ferrenti approach predict a broad range of materials as suitable for quantum technologies. The restrictions set in the intuitive approach, which relies exclusively on findings from the literature, predicted substantially fewer candidates, and provided a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. All four machine learning (ML) methods applied to the training set derived using the intuitive approach agreed on a set of $214$ predicted candidate materials.
All three approaches and all four ML methods agreed on a subset of $47$ eligible candidates %(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for futher material explorations towards quantum technology.

@@

\section*{Introduction}
The second quantum revolution is heralding the development of new solutions and devices based on principles from quantum mechanics \cite{Acin2018}. Among the promising technologies that are already available we find in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum computer architectures and prototypes. 
Quantum computers are in high demand to meet the increasing need for computing power to solve complex many-body problems that are beyond the scope of classical machines.   
Recent advances have highlighted the potential of quantum computer architectures to outperform classical systems as a $53$-qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a $200.000$ core supercomputer \cite{Arute_2019}. Most recently, IBM announced its $127$ qubit quantum processor~\cite{IBM2021}. Simultaneously, the concepts of entanglement and teleportation may enable advanced quantum communication protocols such as quantum cryptography and the quantum internet, spurring further investigations into technologies based on quantum mechanics.

@@

Several platforms are available for the development of quantum technologies (QT), but the materials and fabrication technologies are less mature than those for, e.g., classical computers and sensors. 
An important concern in this context is that of scalability. 
For example, the best performing quantum computer prototypes available today rely on superconducting electronics or trapped ions. Both technologies require millikelvin temperatures to operate, fabrication is challenging, and the stability of interactions between qubits is an issue. Instead, semiconductors are emerging as a promising alternative platform, offering competitive characteristics combined with the possibility of room temperature operation and mature and scalable material processing and fabrication.

Quantum technologies based on semiconductors rely on either defects or quantum dots where the latter kind can be of the self-assembled or fabricated type. 
Semiconductor defects can act as single-photon emitters or spin centers and are compatible with the three main QT categories of computing, communication and sensing.  
These characteristics are most often found for the case of defects that introduce deep energy levels into the semiconductor band gap. So-called deep level defects can trap charge carriers in localized states that are essentially isolated from the surroundings, making them highly suitable for QT due to, e.g., indistinguishable single-photon emission and long spin coherence times. 
The most well-known quantum compatible defect is the negatively charged nitrogen-vacancy (NV) center in diamond \cite{Doherty_2013}, but silicon carbide (SiC) and the various quantum emitters therein are strong contenders in certain areas due to the favorable emission wavelength region in the near infrared coupled with more mature material processing and fabrication \cite{Bathen2021}. 
However, defect based QT is still in the early stages, and the issues left to address include identification of suitable host materials and candidate defects, and scalable and reproducible quantum device fabrication. 
Furthermore, we lack a complete understanding of the requirements for a material to manifest quantum compatible properties,  
and the selection of known quantum compatible semiconductor host materials is slim.

The majority of discoveries of QT compatible characteristics in semiconductors has so far happened by serendipity, and there is an urgent need for a better and more systematic understanding of how the QT compatible materials properties manifest. In this context, a framework for dedicated materials search and analysis is needed.

The fourth science paradigm of big data driven science presents a significant improvement over the third paradigm focusing on computational science, for instance via density functional theory calculations of different materials properties. Instead, the fourth paradigm  introduces the potential of targeted search for promising material systems to enable properties such as coherent and isolated spin states and indistinguishable single-photon emission to manifest. 
Rather than searching through a host of signals for those that match our criteria, we are now able to \textit{predict} which materials and signatures we should target for more detailed studies, following the framework illustrated in Figure~\ref{fig:ht-workflow}. 
This is made feasible by the availability of databases containing material properties for a wide range of different systems. In this work, the data in question are provided by bulk density functional theory (DFT) calculations to obtain the ground state properties of different elements and compounds. Combined with machine learning methods we provide a path towards precise classification of qubit candidates. The inclusion of machine learning methods follows recent trends in applications of statistical learning, data science and machine learning for scientific discoveries, see for example Refs.~\cite{deiana2021,Carleo2019}.

In this work we provide a framework for the data mining and automated discovery of promising solid-state semiconductor hosts for quantum emitters and spin centers using targeted database search and machine learning methods combined with knowledge from the field. 
In this context, we do not distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that a large enough database with relevant theoretical and/or experimental data are available.

The approach we have developed relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
The Python library for data mining Matminer \cite{Ward2018} was then used for material analysis to featurize the extracted data. An important aspect of our work is the database building and pertinent development of the training sets for machine learning methods. We have developed three different approaches to data mining: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) \emph{the augmented Ferrenti approach} and (iii) \emph{the intuitive approach}. We find that the former two methodologies base their data extraction protocols on broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The third approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates. The three training sets where then analyzed with the four supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting, and yielded $214$ predicted candidates 
%(to a $0.5$ cut-off)  
with examples such as ZnGeP$_2$, MgSe, CdS, BP, BC$_2$N, BP, Ge, GeC, InP, and InAs. 
Focused theoretical and experimental studies, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, are needed to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials.

@@

The information stream in this work can be regarded as many modular parts 
%connected in logical pieces. 
%MHJ What do we mean with logical pieces? 
that are connected. 
The initial step for gathering material data and building features is visualized by the  flowchart in Figure~\ref{fig:flowchart}. 
Initially, we start by extracting all entries in the Materials Project (MP) database that match a specific query. 
The MP database contains ground state properties of different materials that are computed using density functional theory (DFT). The DFT calculations in the database were performed using the Vienna ab initio simulation package (VASP) \cite{Kresse1996} and the Perdew-Burke-Ernzerhof (PBE) \cite{Perdew1996} exchange-correlation functional to calculate the electronic structure of the materials.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart.png}
    \caption{The data flow of the main project, starting from an initial MP query, and ending with a featurized dataset with entries from several other databases. To limit the memory and computational usage, the data are partitioned into smaller subsets where the respective Pymatgen objects  
    (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) 
    are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. Abbreviations used are Materials Project (MP), density of states (DOS) and electronic band structure (band).
    }
    \label{fig:flowchart}
\end{figure}

The conditions for the initial MP query are that the materials must derive from 
%experimental data 
the ICSD database and have a band gap wider than $0.1$~eV to exclude metallic compounds. The Inorganic Crystal Structure Database (ICSD) is the world's largest database for completely identified inorganic crystal structures \cite{Allen1987,Zagorac2019}. 
%OH: corrected icsd citation to german one (different from nist)
In a parallel step, entries that are deemed similar to the entries from the initial query are extracted from the databases OQMD \cite{Saal2013,Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015}, AFLOW-ML \cite{Isayev2017} and the Citrination platform \cite{OMaraJordan2016MDIA}. The results of these steps are combined into a data set for further analysis.

Thereafter, we apply Matminer’s \cite{Ward2018} featurization tools to generate thousands of features from the data. 
Before applying any machine learning algorithm, the data was transformed into a numerical representation that reflects the relationship between the input and output data. This transformation is known as generating descriptors 
or features, however, we will in this work adapt the name featurization. 
The open-source library of Matminer provides many tools to featurize existing features extracted from the Materials Project. 
%MB I moved the below to methods 
%To apply Matminer's featurization tools, we extend an existing implementation by \citeauthor{Breuck2021} \cite{Breuck2021}, which were used to generate a supervised machine learning framework called the MODnet. Their implementation provides featurization for a material's composition, structure and atomic sites. However, Matminer also provides featurization tools for a material's density of states (DOS) and band structure, therefore, we extend their implementation to facilitate such featurizations. 
A schematic visualizing the featurization process in Matminer is contained in Fig.~\ref{fig:flowchart} 
and focuses on a material's composition,  structure, atomic sites, density of states and band structure. 
The $39$ features selected as material descriptors in this work can be found in the Supplementary Material at \cite{supplementary}. The selection of features was kept rather wide as we do not precisely know which features best describe a solid state material platform for quantum technology.

@@

After compiling a data set from the procedure outlined above, we face an obstacle in terms of defining a training set for the machine learning algorithms. This is challenging for two reasons.  
Firstly, although QT compatible properties are becoming increasingly well studied for the case of, e.g., nitrogen-vacancy centers in diamond, single-photon emitters in silicon carbide and quantum dot structures, few candidate materials are known. An additional consideration is that the physical mechanisms promoting favorable properties are not fully understood, neither for the case of defect- or QD-based quantum technology. Secondly, defining materials as unsuitable candidates for QT has proven to be surprisingly intricate. 
We follow three separate procedures to define a training set consisting of materials that are labeled as both suitable and unsuitable candidates for QT.

\subsubsection*{The Ferrenti approach}
The first approach to defining a training set is based on the criteria proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}.
They suggest a data mining process consisting of four stages by systematically evaluating the suitability of host materials from the Materials Project. In this framework, we label \textbf{suitable} candidates by the following steps: 
\begin{enumerate}
    \item Include materials that;
    \begin{itemize}
        \item contain elements with more than $50 \ \%$ natural abundance of spin zero isotopes,
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database, 
        and
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Pragmatically remove toxic, radioactive and otherwise ``difficult'' materials;
    \begin{itemize}
        \item exclude Th, U, Cd and Hg because they are radioactive and/or toxic in the most stable forms,
        \item exclude any rare-earth metals (because of the difficulty of obtaining pure materials free of isotopes with nuclear spin) and noble gases (due to the lack of stable solid phases),
        \item exclude transition metal elements with unpaired electrons like Fe and Ni because of their paramagnetism. Ru and Os are also removed because they only exist in the data as complex cluster structures. 
    \end{itemize}
    \item Include only materials with a band gap larger than $0.5$~eV calculated using DFT and the PBE functional. The value of $0.5$~eV was chosen to match that typically predicted for silicon by PBE-level DFT calculations. 
    \item Ensure that energy above hull is less than $0.2$~eV per atom.
\end{enumerate}

The inclusion criteria are targeted primarily at semiconductors that can host deep level defects with spin qubit capabilities. In this context, long spin coherence times are needed, and the possibility of having an environment free of nuclear spins and permanent magnetism is therefore desirable. Additionally, to obtain defect-based single-photon emitters with sharp zero-phonon lines and indistinguishable photon emission, non-polar materials are assumed to be preferable. 
The inclusion criteria listed above are based on the work of \citeauthor{Weber2010} \cite{Weber2010}.

Transition metal elements are eliminated in this approach if they have unpaired electrons because the presence of permanent electric dipole moments may have a detrimental impact on the optical coherence of defect emission. In this context, also the symmetry of the crystalline host may be important. Therefore, only phases which take on non polar space groups are considered in this approach. 
The energy above hull requirement is included to ensure that the selected compounds are thermodynamically stable.

We then label \textbf{unsuitable} candidates according to the reverse requirements of the above; as materials in the ICSD database \cite{Allen1987,Zagorac2019} %MHJ need ref here
that crystallize in polar space groups, are calculated to be magnetic and have a band gap larger than $0.1$~eV (to exclude metals but include lower-band gap semiconductors) as calculated by DFT using the PBE functional, as discussed above.
The resulting complete training set contains $1581$ materials where %$56 \ \%$ 
$35 \ \%$
are labeled as unsuitable and 
%$44 \ \%$
$65 \ \%$
as suitable for QT applications.

\subsubsection*{The augmented Ferrenti approach}
In the second approach, we adjust the former one in order to expand the training set beyond practical considerations. This approach is therefore named \emph{the augmented Ferrenti approach}.

In the first approach we included criteria that were not physically motivated, but instead to some extent based on a practical perspective, such as excluding toxic and radioactive elements. Such criteria are not necessarily related to what makes a material suitable for quantum technology - eventually it is up to the experimentalist to decide the feasibility of the material to be employed. 
Consequently, in the augmented Ferrenti approach, we remove stage two from the approach above. Moreover, we include a few additional elements that have shown promising properties but were initially excluded due to the lack of spin zero isotopes.

The following steps constitute the process of choosing \textbf{suitable} candidates in the \emph{augmented Ferrenti approach}:
\begin{enumerate}
    \item Include materials that; 
    \begin{itemize}
        \item contain elements where more than half have a natural abundance of spin zero isotopes, including Al, P, Ga, As, B and N, 
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database,
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Only keep materials that have a band gap larger than $1.5$~eV as calculated by PBE-GGA. The higher band gap requirement (as compared to the Ferrenti approach) is included here to avoid including an unfeasible number of candidate materials. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom. 
\end{enumerate}

For \textbf{unsuitable} candidates, we implement the same strategy as for the first Ferrenti approach. The result is a somewhat unbalanced training set, with up to $75 \ \%$ of the materials found in the suitable group. However, the training set is $78 \ \%$ larger than for the Ferrenti approach with $2141$ suitable and $684$ unsuitable materials.

>>>>

\begin{abstract}
Semiconductor materials provide a compelling platform for quantum technologies (QT), and the properties of a vast amount of materials can be found in databases containing information from both experimental and theoretical explorations. However, scouring these databases to find promising candidate materials for quantum technology applications among tens of thousands of contenders is a major challenge. Therefore, we have developed a framework for the automated discovery of semiconductor host platforms for QT using material informatics and machine learning methods.

Data extraction tools have been developed for a number of materials science databases resulting in a data set consisting of over $25.000$ materials. The information from the databases was converted into nearly $5000$ physics-informed features. Furthermore, we have developed and implemented three different approaches for the data mining process: \textit{The Ferrenti approach}, \textit{the extended Ferrenti approach}, and \textit{the empirical approach}. 
The three approaches define distinct training sets for the supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting.

We find that the Ferrenti approach and the extended Ferrenti approach predict a broad range of materials as suitable for quantum technologies. The restrictions set in the empirical approach, which relies exclusively on findings from the literature, predicted substantially fewer candidates, and provided a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. All four machine learning (ML) methods applied to the training set derived using the empirical approach agreed on a set of $214$ predicted candidate materials.
All three approaches and all four ML methods agreed on a subset of $47$ eligible candidates %(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for futher material explorations towards quantum technology.

@@

\section*{Introduction}
Quantum technologies (QT) based on solid state platforms are rapidly gaining in both success and notoriety. 
Among the promising applications that are already available we find important breakthroughs  such as in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum information processing prototypes and architectures.  
Quantum computers are in high demand to meet the increasing need for computing power to solve complex many-body problems that are beyond the scope of classical machines.   
Recent advances have highlighted the potential of quantum information processors to outperform classical systems. Indeed, Google's $53$ qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a $200.000$ core supercomputer \cite{Arute_2019}. Most recently, IBM announced its $127$ qubit quantum processor~\cite{IBM2021}. Simultaneously, the concepts of entanglement and teleportation may eventually facilitate advanced quantum communication protocols such as quantum cryptography and the quantum internet, spurring further investigations into technologies based on quantum mechanics.

@@

Several platforms are available for the development of quantum technologies, but the materials and fabrication technologies are less mature than those for, e.g., classical computers and sensors. 
An important concern in this context is that of scalability. 
For example, the best performing quantum computer prototypes available today rely on superconducting electronics, requiring millikelvin temperatures to operate, with challenging fabrication and the stability of interactions between qubits is an issue. Instead, semiconductors are emerging as a promising alternative platform, offering competitive characteristics combined with the possibility of room temperature operation and mature and scalable material processing and fabrication.

Quantum technologies based on semiconductors rely on either defects or quantum dots where the latter kind can be of the self-assembled or fabricated type. 
Semiconductor defects can act as single-photon emitters or spin centers and are compatible with the three main QT categories of computing, communication and sensing.  
These characteristics are most often found for the case of defects that introduce deep energy levels into the semiconductor band gap. So-called deep level defects can trap charge carriers in localized states that are essentially isolated from the surroundings, making them highly suitable for QT due to, e.g., indistinguishable single-photon emission and long spin coherence times. 
The most well-known quantum compatible defect is the negatively charged nitrogen-vacancy (NV) center in diamond \cite{Doherty_2013}, but silicon carbide (SiC) and the various quantum emitters therein are strong contenders especially for quantum communication purposes due to the favorable emission wavelength region in the near infrared coupled with more mature material processing and fabrication \cite{Bathen2021}. 
However, defect based QT is still in the early stages, and the issues left to address include identification of suitable host materials and candidate defects, and scalable and reproducible quantum device fabrication. 
Furthermore, we lack a complete understanding of the requirements for a semiconductor material to manifest quantum compatible properties,  
and the selection of known quantum compatible host materials is slim.

The majority of discoveries of QT compatible characteristics in semiconductors has so far happened by serendipity, and there is an urgent need for a better and more systematic understanding of which material requirements must be met for QT compatible characteristics (e.g. single-photon emission and manipulable spin states) to manifest. In this context, a framework for dedicated materials search and analysis is needed.

The fourth science paradigm of big data driven science presents a significant improvement over the third paradigm focusing on computational science, for instance via density functional theory calculations of different materials properties. Instead, the fourth paradigm introduces the potential of targeted search for promising material systems in which to expect QT compatible properties. 
Rather than searching through a host of signals for those that match our criteria, we are now able to \textit{predict} which materials and signatures we should target for more detailed studies, following the framework illustrated in Figure~\ref{fig:ht-workflow}. 
This is made feasible by the availability of databases containing material properties for a wide range of different systems. In this work, the data in question are provided by bulk density functional theory (DFT) calculations to obtain the ground state properties of different elements and compounds. Combined with machine learning methods we provide a path towards precise classification of qubit candidates. The inclusion of machine learning methods follows recent trends in applications of statistical learning, data science and machine learning for scientific discoveries, see for example Refs.~\cite{deiana2021,Carleo2019}.

In this work we provide a framework for the data mining and automated discovery of promising solid-state semiconductor hosts for quantum emitters and spin centers using targeted database search and machine learning methods combined with knowledge from the field. 
In this context, we do not distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that high quality databases containing relevant theoretical and/or experimental data are available.

The approach we have developed relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
The Matminer Python library for data mining \cite{Ward2018} was then used for material analysis to featurize the extracted data. An important aspect of our work is the database building and pertinent development of the training sets for machine learning methods. We have developed three different approaches to data mining: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) 
%Marianne I have now changed augmented Ferrenti to extended Ferrenti 
\emph{the extended Ferrenti approach} and (iii) \emph{the empirical approach}. We find that the former two methodologies base their data extraction protocols on broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The third approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates. The three training sets where then analyzed with the four supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting, and yielded $214$ predicted candidates that were common between all approaches and ML methods. 
%(to a $0.5$ cut-off)  
Example materials include ZnGeP$_2$, MgSe, CdS, BP, BC$_2$N, BP, Ge, GeC, InP, and InAs. 
Focused theoretical and experimental studies, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, are needed to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials.

@@

The information stream in this work can be regarded as many modular parts 
%connected in logical pieces. 
%MHJ What do we mean with logical pieces? 
that are connected. 
The initial step for gathering material data and building features is visualized by the  flowchart in Figure~\ref{fig:flowchart}. 
Initially, we start by extracting all entries in the Materials Project (MP) database that match a specific query. 
The MP database contains ground state properties of different materials that are computed using density functional theory (DFT). The DFT calculations in the database were performed using the Vienna ab initio simulation package (VASP) \cite{Kresse1996} and the Perdew-Burke-Ernzerhof (PBE) \cite{Perdew1996} exchange-correlation functional to calculate the electronic structure of the materials. 
We note that despite being immensely successful in describing a number of material properties, PBE is widely known for underestimating the band gap of semiconductors \cite{Freysoldt2014}. Therefore, not all properties predicted using the PBE functional are reliable, and the band gaps in particular cannot be trusted in absolute terms. Nonetheless, the functional is in wide use due to the combination of reasonable accuracy and high computational throughput, and is usually considered to be reliable for large-scale trends in semiconductor properties.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart_2.png}
    \caption{The data flow of the main project, starting from an initial MP query, and ending with a featurized data set with entries from several other databases. To limit the memory and computational usage, the data are partitioned into smaller subsets where the respective Pymatgen objects  
    (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) 
    are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. Abbreviations used are Materials Project (MP), density of states (DOS) and electronic band structure (band).
    }
    \label{fig:flowchart}
\end{figure}

The conditions for the initial MP query are that the materials must derive from 
%experimental data 
the Inorganic Crystal Structure Database (ICSD) database and have a band gap wider than $0.1$~eV to exclude metallic compounds. The ICSD is the world's largest database for completely identified inorganic crystal structures \cite{Allen1987,Zagorac2019}. 
%OH: corrected icsd citation to german one (different from nist)
In a parallel step, entries that are deemed similar to the entries from the initial query are extracted from the databases OQMD \cite{Saal2013,Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015}, AFLOW-ML \cite{Isayev2017} and the Citrination platform \cite{OMaraJordan2016MDIA}. The results of these steps are combined into a data set for further analysis.

Thereafter, we apply Matminer’s \cite{Ward2018} tools to generate thousands of features from the data. 
Before applying any machine learning algorithm, the data was transformed into a numerical representation that reflects the relationship between the input and output data. This transformation is known as generating descriptors 
or features, however, we will in this work adopt the name featurization. 
The open-source library of Matminer provides many tools to featurize existing features extracted from the Materials Project. 
%MB I moved the below to methods 
%To apply Matminer's featurization tools, we extend an existing implementation by \citeauthor{Breuck2021} \cite{Breuck2021}, which were used to generate a supervised machine learning framework called the MODnet. Their implementation provides featurization for a material's composition, structure and atomic sites. However, Matminer also provides featurization tools for a material's density of states (DOS) and band structure, therefore, we extend their implementation to facilitate such featurizations. 
A schematic visualizing the featurization process in Matminer is contained in Fig.~\ref{fig:flowchart} 
and focuses on a material's composition,  structure, atomic sites, density of states and band structure. 
The $39$ featurizers selected as material descriptors in this work are summarized and described in the Supplementary Material at \cite{supplementary}. The selection of features was kept rather wide as we do not precisely know which features best describe a solid state material platform for quantum technology.

%It is important to note that 
The constructed data sets encompass compounds formed by a plethora of combinations of surfaces, interfaces, nanostructures, compositions and structures, but this complexity is not necessarily reflected in the material descriptors. 
Furthermore, we have utilized data obtained from high-throughput density functional theory calculations. Indeed, there are possible errors associated with every step, starting from an initial calculation, adding of data in the database, gathering of data, featurization of data, preprocessing of data, data mining, and finally training a model and making a prediction. Unfortunately, if an error has happened in the first part of the process, the error follows the entire process and will get increasingly harder to detect. Therefore, we are dependent on that the Materials Project database contains data with high quality, but we note that some errors are likely present in our data set.

@@

After compiling a data set from the procedure outlined above, we must define a training set for the machine learning algorithms. This is challenging for two reasons.  
Firstly, although QT compatible properties are becoming increasingly well studied for the case of, e.g., nitrogen-vacancy centers in diamond, single-photon emitters in silicon carbide and quantum dot structures, relatively few candidate materials are known to be suitable. An additional consideration is that the physical mechanisms promoting favorable properties are not fully understood, neither for the case of defect- or QD-based quantum technologies. Secondly, defining materials as unsuitable candidates for QT has proven to be surprisingly intricate. 
We follow three separate procedures for defining training sets consisting of materials that are labeled as both suitable and unsuitable candidates for QT.

\subsubsection*{The Ferrenti approach}
The first approach to defining a training set is based on the criteria proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}.
They suggest a data mining process consisting of four stages by systematically evaluating the suitability of host materials from the Materials Project. In this framework, we label \textbf{suitable} candidates by the following steps: 
\begin{enumerate}
    \item Include materials that;
    \begin{itemize}
        \item contain elements with more than $50 \ \%$ natural abundance of spin zero isotopes,
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database, 
        and
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Pragmatically remove toxic, radioactive and otherwise ``difficult'' materials;
    \begin{itemize}
        \item exclude Th, U, Cd and Hg because they are radioactive and/or toxic in the most stable forms,
        \item exclude any rare-earth metals (because of the difficulty of obtaining pure materials free of isotopes with nuclear spin) and noble gases (due to the lack of stable solid phases),
        \item exclude transition metal elements with unpaired electrons like Fe and Ni because of their paramagnetism, Ru and Os are also removed because they only exist in the data as complex cluster structures. 
    \end{itemize}
    \item Include only materials with a band gap larger than $0.5$~eV calculated using DFT and the PBE functional. The value of $0.5$~eV was chosen to match that typically predicted for silicon by PBE-level DFT calculations. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom.
\end{enumerate}

These inclusion criteria are based on the work of \citeauthor{Weber2010} \cite{Weber2010} and targeted primarily at semiconductors that can host deep level defects with spin qubit capabilities. In this context, long spin coherence times are needed, and the possibility of having an environment free of nuclear spins and permanent magnetism is therefore desirable. Additionally, to obtain defect-based single-photon emitters with sharp zero-phonon lines and indistinguishable single-photon photon emission, non-polar materials are assumed to be preferable. 
%The inclusion criteria listed above are based on the work of \citeauthor{Weber2010} \cite{Weber2010}.

Transition metal elements are eliminated in this approach if they have unpaired electrons because the presence of permanent electric dipole moments may have a detrimental impact on the optical coherence of defect emission. In this context, also the symmetry of the crystalline host may be important. Therefore, only phases which take on non-polar space groups are considered in this approach. 
The energy above hull requirement is included to ensure that the selected compounds are thermodynamically stable.

Next, \textbf{unsuitable} candidates are labeled according to the reverse requirements of the above; as materials in the ICSD database \cite{Allen1987,Zagorac2019} 
that crystallize in polar space groups, are calculated to be magnetic and have a band gap larger than $0.1$~eV in the MP database (to exclude metals but include lower-band gap semiconductors). 
The resulting complete training set contains $1581$ materials where $35 \ \%$
are labeled as unsuitable and 
$65 \ \%$ as suitable for QT applications.

%The training set from the Ferrenti approach reveal that carbon in a diamond-like structure is classified as a suitable candidate, in good agreement with experimental observations, (see Table~\ref{tab:qt-materials}) and hence also with the candidates that were manually defined as suitable in the empirical approach.  Interestingly, in the training set from the Ferrenti approach, we find that carbon in two-dimensional graphite-like structures are labeled as suitable as well. All structures of silicon are also labeled as suitable together with one entry of silicon carbide. Note that this is the 3C polytype of SiC, meaning that the most well-established quantum compatible SiC polytype, 4H-SiC, is missing from the training set. Among other potentially suitable candidates we find ZnS, ZnSe, ZnO and ZnTe present in the training data.

\subsubsection*{The extended Ferrenti approach}
In the second approach, we adjust the former one in order to expand the training set beyond practical considerations. This approach is therefore named \emph{the extended Ferrenti approach}.

In the first approach we included criteria that were 
%not physically motivated, but instead 
to some extent based on a practical perspective, such as excluding toxic and radioactive elements. Such criteria are not necessarily related to what makes a material suitable for quantum technology - eventually it is up to the experimentalist to decide the feasibility of the material to be employed. 
Consequently, in the extended Ferrenti approach, we remove stage two from the approach above. Moreover, we include a few additional elements that have shown promising properties but were initially excluded due to the lack of spin zero isotopes.

The following steps constitute the process of choosing \textbf{suitable} candidates in the \emph{extended Ferrenti approach}:
\begin{enumerate}
    \item Include materials that; 
    \begin{itemize}
        \item contain elements where more than half have a natural abundance of spin zero isotopes, including Al, P, Ga, As, B and N, 
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database,
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Only keep materials that have a band gap larger than $1.5$~eV in the MP database. The higher band gap requirement (as compared to the Ferrenti approach) is included here to avoid including an unfeasible number of candidate materials in the training set. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom. 
\end{enumerate}

For \textbf{unsuitable} candidates, we implement the same strategy as for the first Ferrenti approach. The result is a somewhat unbalanced training set, with up to $75 \ \%$ of the materials found in the suitable group. However, the training set is $78 \ \%$ larger than for the Ferrenti approach consisting of  $2141$ suitable and $684$ unsuitable materials.

%MB I moved this down to not confuse the order of information 
% 
% In the training set of the extended Ferrenti approach, we find a single entry for each of SiC, Si, GaN, ZnS, GaP, AlAs and AlP, carbon in both diamond- and graphite-like structures, and AlN in three different configurations. The training set includes a larger variety of materials that are known to be quantum compatible as compared to the Ferrenti approach due to admitting more elements in the initial selection process. However, since we also included a more stringent band gap restriction of $1.5$~eV (to restrict the amount of materials in the training set), we find a more sparse representation of each known chemical formula present in the training set.
